---
title: "Data607-Assignment2"
author: "Santosh Manjrekar"
date: "September 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("RMySQL")
library(RMySQL)
```

Now to make the connection. With some tests to make sure the connection works.

```{r}
movie_db = dbConnect(MySQL(), user='root', password='root', dbname='moviedb', host='localhost')
dbListTables(movie_db)
dbListFields(movie_db, 'movies')
```

I sourced my data on Facebook using the movies( Black Panther,Avengers-Infinity War,Incredible2,Deadpool2,Annihilation,Mission_impossible).
My friends are a mix of geeks and/or parents of young kids so I included movies likely to get a response. I recieved more then required by the assignment within minutes.

now to Read the database into the dataframe:
```{r}
movie_df <- dbReadTable(movie_db, name = "movies")
movie_df <- data.frame(movie_df)
movie_df
```

While we are at it let's take a look at the data, keeping in mind that 0 means they did not see the film. Since I did have friends that insisted in half scores, histograms work a little better than bar plots in R even though the data is still discrete.

```{r}
hist(movie_df$black_panther)
hist(movie_df$avengers_iw)
hist(movie_df$incredible2)
hist(movie_df$deadpool2)
hist(movie_df$annihilation)
hist(movie_df$mission_impo)
```

####Conclusion:

It looks like Incredible2 was the most liked movie with Mission Impossible coming in a close second, Deadpool2 was the least liked an annihilation had the most spread in opinion.